Moral Foundations of U.S. Political News Organizations

### 2018-12-22 pilot_manuscript.Rmd
Added moral psych overview to intro

Something clever to start

@@

@@

This issue is compounded by the fact the Moral Foundations Questionnaire (MFQ) and its subscales assume endorsement lies on a continuum. The Moral Foundations Questionnaire (MFQ) measures individuals’ endorsements of each of the foundations using a six-point scale [@Graham2011]. The quesionnaire is made up of judgment items and relevance items. Judgment items are phrased such that the respondent signals their agreement or disagreement with straightforward statements. An example of such a statement reads: “It can never be right to kill another human being” [@Graham2011]. Relevance items gauge the respondent's opinion regarding the importance of foundation-related concerns. For example, the respondent is directed to rate how important the following situation is to their sense of morals: "whether or not someone did something disgusting." This example measures the relevance of the *purity/sanctity* foundations. Each foundation has a judgment and relevance subscale, totalling 10 subscales for the MFQ [@Graham2011].

>>>>

Morality is prevalent/important

Domain of philosophers

Then psychologists (more on that later)

Btw, it's intrinsically linked with politics

Lots of partisan media choices today = different appeals to audience's moral concerns?

@@

As a discipline, modern moral psychology started in the late 1960s with Lawrence Kohlberg [@Haidt2007]. Kohlberg's research popularized his theory of the devlopment of moral reasoning. This theory establishes the steps of moral reasoning through which humans proceed as they mature. Kohlberg borrwowed from Jean Piaget's stages of cognitive devlopment in which children progress from the sensorimotor through to the formal operations stage [@Haidt2007]. Similarly, Kohlberg found people typcially start with a "pre-conventional" understanding of morality during infancy and progress through several steps until they reach a "post-conventional" ethics [@Haidt2007]. People who have reached the post-conventional stage are said to be able to weigh competing abstractions and reason their way to a conclusion that promotes justice. From Kohlberg's perspective, issues of justice and fairness comprise the foundation of morality [@Haidt2007]. This view persisted until it encountered criticism in the early 1980s.

Kohlberg's conception of morality faced major scrutiny from psycholgist Carol Gilligan. In 1982, Gilligan criticized Kohlberg's theory on the grounds that it focused solely on the moral concerns of men, and that it ignored those of women [@Haidt2007]. She offered a historic argument contending women have traditionally filled roles related to caring and nurturing. Gilligan thus asserted morality was built upon an alternative moral foundation: caring [@Haidt2007]. This debate between competing conceptions of morality did not resolve until Gilligan and Kohlberg conceded the existence of two moral foundations: justice and caring [@Haidt2007]. While this new direction in moral psychology appeared to represent a more inclusive outlook on the construct, these novel ideas would soon be challenged on the grounds of its apparent western-centric outlook.

In their brief overview of the history of moral psychology, @Graham2009 explained Shweder, Much, Mahapatra, and Park's objections to moral psychology as it stood in the late 1980s. Their criticism centered on the fact moral psychology concerned itself with issues regarding justice and individuals' rights. Such a system, they argued, did not account for moral concerns outside of the western world [@Graham2009]. Individually focused concerns can be grouped under an overarching "ethic of autonomy," which was thought to be one of three ethics upon which humans base moral decisions. The other two ethics were the "ethic of community" (comprising one's duty to their family, tribe, etc.) and the "ethic of divinity"" - representing one's duty not to defile their God-given body and soul [@Graham2009]. In the 2000s, @Haidt2007 took this line of reasoning further in their assertion that moral psychology favored certain political ideologies over others.

@@

This issue is compounded by the fact the Moral Foundations Questionnaire (MFQ) and its subscales assume endorsement lies on a continuum. The Moral Foundations Questionnaire (MFQ), which was developed subsequent to the MFD, measures individuals’ endorsements of each of the foundations using a six-point scale [@Graham2011]. The quesionnaire is made up of judgment items and relevance items. Judgment items are phrased such that the respondent signals their agreement or disagreement with straightforward statements. An example of such a statement reads: “It can never be right to kill another human being” [@Graham2011]. Relevance items gauge the respondent's opinion regarding the importance of foundation-related concerns. For example, the respondent is directed to rate how important the following situation is to their sense of morals: "whether or not someone did something disgusting." This example measures the relevance of the *purity/sanctity* foundations. Each foundation has a judgment and relevance subscale, totalling 10 subscales for the MFQ [@Graham2011].
